### 2024-03-16 paper/paper.qmd
Cleaning instructions added in appendix

---
title: "Biden Projected to Win the 2024 US Presidential Elections"
subtitle: "Analysis using data from American Community Survey 2022 and Ipsos Poll 2023 levereging Bayesian Logistic Regression with Post-Stratification"
author:
  - Monica Sainani
  - Kaavya Kalani
thanks: "Code and data supporting this analysis is available at: https://github.com/sainanim/Predictions-for-the-US-2024-Elections"
date: today
date-format: long
abstract: "We analyze American voting behavior, focusing on demographic, socioeconomic, and geographic factors. We use a Bayesian Logistic Regression model with Post-Stratification to predict the election outcome between Biden and Trump. We find a strong relationship between age, family income, education, state of residency, sex and likely voting outcome. Our central forecast is a Biden re-election. This research provides insights into the factors affecting political engagement in the U.S., potentially guiding future electoral strategies and policy discussions."
format: pdf
number-sections: true
toc: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

# install.packages("broom.mixed")
# install.packages("tidyverse")
# install.packages("arrow")
# install.packages("dplyr")
# install.packages("knitr")
# install.packages("ggplot2")
# install.packages("modelsummary")
# install.packages("here")
# install.packages("kableExtra")

library(broom.mixed)
library(tidyverse)
library(arrow)
library(dplyr)
library(knitr)
library(ggplot2)
library(modelsummary)
library(here)
library(kableExtra)

# Read the datasets and model
poststrat_data <- arrow::read_parquet(file = here("data/analysis_data/poststrat_data.parquet"))
survey_data <- arrow::read_parquet(file = here("data/analysis_data/survey_data.parquet"))
us_election_model <- readRDS(file = here("models/single_bay.rds"))

# Read the post-stratified data
state_vote_for_biden <- read.csv(here("data/poststratified/state_vote_for_biden.csv"))
age_vote_for_biden <- read.csv(here("data/poststratified/age_vote_for_biden.csv"))
sex_vote_for_biden <- read.csv(here("data/poststratified/sex_vote_for_biden.csv"))
general_data <- read.csv(here("data/poststratified/general_vote_for_biden.csv"))

```

# Introduction

# Data

For this analysis, we have used two datasets. The datasets were cleaned and analysed using the statistical programming software R [@citeR] along with the help tidyverse [@citeTidyverse], knitr [@citeKnitr], ggplot2 [@citeGgplot], here [@citeHere], dplyr [@citeDplyr], arrow [@citeArrow], rstanarm [@citeRstanarm], broom.mixed [@citeBroomMixed], modelsummary [@citeModelSummary] and kableExtra [@citeKableExtra].

## Survey Data

Our survey data is from the Ipsos Poll [Roper #31120637]. It was collated from a web-based survey conducted on November 3 - 4, 2023 using KnowledgePanel. The survey is conducted by Ipsos and the sponsor is ABC News. This poll is based on a nationally representative probability sample of 949 adults age 18 or older with oversamples among Black and Hispanic respondents. The poll is a collection of demographic information and political preferences.

A person becomes an entry in this dataset if they were 18 years or older on November 3-4, 2023, an American national and participated in this poll. They are recruited through a scientifically developed addressed-based sampling methodology using the latest Delivery Sequence File of the USPS – a database with full coverage of all delivery points in the US. Households invited to join the panel are randomly selected from all available households in the US. People in the sampled households are invited to join and participate in the panel. Those selected who do not already have internet access are provided a tablet and internet connection at no cost to the panel member. Those who join the panel and who are selected to participate in a survey are sent a unique password-protected log-in used to complete surveys online. Panelists receive a unique login to the survey and are only able to complete it one time. No reminder emails were sent for this study.

The results from this dataset are used for our model. Among the range of variables available, we selected age, income, education, state and sex as our independent variables.

- Age is the age group the panelist falls into. The poll had the ages of the panelists at the time of the poll which we then classified into one of the four age groups: 18-29, 30-44, 45-59, 60+

- Income is the annual household income reported by the panelist. They were given the option of seven categories: Less than $10,000, $10,000 to $24,999, $25,000 to $49,999, $50,000 to $74,999, $75,000 to $99,999, $100,000 to $149,999 and $150,000 or more.

- Education is the highest level of education attained by the panelist. They were given the option between: Less than high school, High school, Some college and Bachelor's degree or higher. We then classified these into three categories:  Less than high school, High school,  More than high school

- State is the state of residence of the panelist.

- Sex is the sex reported by the panelist and it is either male or female

For the variable we are predicting, we created a binary variable which is 1 if the panelist favors Biden and 0 if the panelist favors Trump. There was no question in the poll which asked this direct question so we used “Overall, do you have a favorable or unfavorable impression of...Donald Trump?”. We acknowledge that not favoring Donald Trump does not necessarily imply favoring Joe Biden, but since they are the two major potential candidates, we classify the panelists who chose “Unfavorable” as supporting Biden, i.e not supporting Trump.

```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-cap: Vote count distribution in survey data
#| label: fig-vote-biden

ggplot(survey_data, aes(x = as.factor(vote_biden))) +
  geom_bar(fill = "skyblue", color = "black") +
  labs(x = "Party Nominee",
       y = "Count") + 
  scale_x_discrete(labels = c("Vote for Trump", "Vote for Biden"))
```

@fig-vote-biden shows the distribution of vote count in our survey data. We see that the survey panelists are heavily in favor of Joe Biden with a over twice of the panelists not favoring Trump. The bias can be a potential weakness of the dataset and we aim to uncover the factors influencing these political preferences.

```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-cap: Variables of interest and the vote proportion distribution
#| fig-subcap: ["Age", "Sex", "Income", "Education", "State"]
#| label: fig-variables
#| layout-ncol: 2

# Calculate the proportion of individuals who voted for Biden for each age
age_proportions <- survey_data %>%
  group_by(age, vote_biden) %>%
  summarize(count = n()) %>%
  mutate(proportion = count / sum(count))

ggplot(age_proportions, aes(x = age, y = proportion, color = as.factor(vote_biden))) +
  geom_point(position = position_dodge(width = 0.5), size = 3) +
  geom_line(aes(group = vote_biden), position = position_dodge(width = 0.5)) +
  labs(x = "Age Category", y = "Proportion", color = "Vote") +
  scale_color_manual(values = c("blue", "red"), labels = c("Vote for Trump", "Vote for Biden")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Calculate the proportion of individuals who voted for Biden for each sex
sex_proportions <- survey_data %>%
  group_by(sex, vote_biden) %>%
  summarize(count = n()) %>%
  mutate(proportion = count / sum(count))

ggplot(sex_proportions, aes(x = sex, y = proportion, color = as.factor(vote_biden))) +
  geom_point(position = position_dodge(width = 0.5), size = 3) +
  geom_line(aes(group = vote_biden), position = position_dodge(width = 0.5)) +
  labs(x = "Sex", y = "Proportion", color = "Vote") +
  scale_color_manual(values = c("blue", "red"), labels = c("Vote for Trump", "Vote for Biden")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Calculate the proportion of individuals who voted for Biden for each income level
income_proportions <- survey_data %>%
  group_by(income, vote_biden) %>%
  summarize(count = n()) %>%
  mutate(proportion = count / sum(count))

ggplot(income_proportions, aes(x = income, y = proportion, color = as.factor(vote_biden))) +
  geom_point(position = position_dodge(width = 0.5), size = 3) +
  geom_line(aes(group = vote_biden), position = position_dodge(width = 0.5)) +
  labs(x = "Income", y = "Proportion", color = "Vote") +
  scale_color_manual(values = c("blue", "red"), labels = c("Vote for Trump", "Vote for Biden")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Calculate the proportion of individuals who voted for Biden for each education level
education_proportions <- survey_data %>%
  group_by(education, vote_biden) %>%
  summarize(count = n()) %>%
  mutate(proportion = count / sum(count))

ggplot(education_proportions, aes(x = education, y = proportion, color = as.factor(vote_biden))) +
  geom_point(position = position_dodge(width = 0.5), size = 3) +
  geom_line(aes(group = vote_biden), position = position_dodge(width = 0.5)) +
  labs(x = "Education", y = "Proportion", color = "Vote") +
  scale_color_manual(values = c("blue", "red"), labels = c("Vote for Trump", "Vote for Biden")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Calculate the proportion of individuals who voted for Biden for each state
state_proportions <- survey_data %>%
  group_by(state, vote_biden) %>%
  summarize(count = n()) %>%
  mutate(proportion = count / sum(count))

ggplot(state_proportions, aes(x = state, y = proportion, color = as.factor(vote_biden))) +
  geom_point(position = position_dodge(width = 0.5), size = 3) +
  geom_line(aes(group = vote_biden), position = position_dodge(width = 0.5)) +
  labs(x = "State", y = "Proportion", color = "Vote") +
  scale_color_manual(values = c("blue", "red"), labels = c("Vote for Trump", "Vote for Biden")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```
@fig-variables shows the vote proportion distribution for our chosen variables of interests. We notice a noticeable difference in the proportions that favor Trump or Biden. Some striking trends we notice in our survey data are more male panelists supporting Trump, majority of Trump supporters being in the 45-59 age group and Trump support declining with increase in education levels. While most states seem to support Trump, we see that some states support Trump over Biden. This is a fresh change to see over all other factors showing a clear preference for Biden in all categories.

## Post-stratification Data

Our post-stratification data is from the most recent American Community Survey (ACS), the 2022 Sample. We use this to represent the voter population and adjust the weight of the survey data. The population for this sample was the entire US population, categorized into households as subgroups, and the responses of the sampled subgroups were collected. ACS 2022 dataset consists of 3,373,378 observations. Using a 1-in-100 national random sample, the ACS 2022 sample provides a thorough overview of the demographic, socioeconomic, and geographic features of the US population. Most importantly, this dataset includes people who live in both home and group quarters, offering a comprehensive view of the American population. A person becomes an entry in this dataset if they are a part of the sampled subgroups and have responded to the survey.

There are a large number of different features available and the variables we chose are age, income, education, state and sex to match with the variables from our survey. These variables were cleaned and the dataset was preprocessed so the values and categories match with the survey dataset.

```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-cap: Comparison of Survey and Census Data Proportions
#| fig-subcap: ["Age", "Sex", "Income", "Education", "State"]
#| label: fig-proportions
#| layout-ncol: 2

library(dplyr)
library(ggplot2)
library(reshape2)

# Group and summarize survey_data by age
df_survey <- survey_data %>% 
  count(age) %>%
  mutate(sur_prop = n / sum(n)) %>%
  select(age, sur_prop)

# Group and summarize poststrat_data (census data) by age
df_census <- poststrat_data %>% 
  count(age) %>% 
  mutate(census_prop = n / sum(n)) %>%
  select(age, census_prop)

# Merge the two datasets
combined_data <- left_join(df_census, df_survey, by = "age")

# Plotting
ggplot(combined_data, aes(age)) +
  geom_point(aes(y = census_prop, color = "Census data"), size = 3) +
  geom_point(aes(y = sur_prop, color = "Survey data"), size = 3) +
  scale_color_manual(values = c("Census data" = "salmon", "Survey data" = "skyblue")) +
  labs(title = "Age",
       x = "Age",
       y = "Proportion") +
  theme_minimal()

# Group and summarize survey_data by sex
df_survey <- survey_data %>% 
  count(sex) %>%
  mutate(sur_prop = n / sum(n)) %>%
  select(sex, sur_prop)

# Group and summarize poststrat_data (census data) by sex
df_census <- poststrat_data %>% 
  count(sex) %>% 
  mutate(census_prop = n / sum(n)) %>%
  select(sex, census_prop)

# Merge the two datasets
combined_data <- left_join(df_census, df_survey, by = "sex")

# Plotting
ggplot(combined_data, aes(sex)) +
  geom_point(aes(y = census_prop, color = "Census data"), size = 3) +
  geom_point(aes(y = sur_prop, color = "Survey data"), size = 3) +
  scale_color_manual(values = c("Census data" = "salmon", "Survey data" = "skyblue")) +
  labs(title = "Sex",
       x = "Sex",
       y = "Proportion") +
  theme_minimal()

# Group and summarize survey_data by income
df_survey <- survey_data %>% 
  count(income) %>%
  mutate(sur_prop = n / sum(n)) %>%
  select(income, sur_prop)

# Group and summarize poststrat_data (census data) by income
df_census <- poststrat_data %>% 
  count(income) %>% 
  mutate(census_prop = n / sum(n)) %>%
  select(income, census_prop)

# Merge the two datasets
combined_data <- left_join(df_census, df_survey, by = "income")

# Plotting
ggplot(combined_data, aes(income)) +
  geom_point(aes(y = census_prop, color = "Census data"), size = 3) +
  geom_point(aes(y = sur_prop, color = "Survey data"), size = 3) +
  scale_color_manual(values = c("Census data" = "salmon", "Survey data" = "skyblue")) +
  labs(title = "Income",
       x = "Income",
       y = "Proportion") +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

# Group and summarize survey_data by education
df_survey <- survey_data %>% 
  count(education) %>%
  mutate(sur_prop = n / sum(n)) %>%
  select(education, sur_prop)

# Group and summarize poststrat_data (census data) by education
df_census <- poststrat_data %>% 
  count(education) %>% 
  mutate(census_prop = n / sum(n)) %>%
  select(education, census_prop)

# Merge the two datasets
combined_data <- left_join(df_census, df_survey, by = "education")

# Plotting
ggplot(combined_data, aes(education)) +
  geom_point(aes(y = census_prop, color = "Census data"), size = 3) +
  geom_point(aes(y = sur_prop, color = "Survey data"), size = 3) +
  scale_color_manual(values = c("Census data" = "salmon", "Survey data" = "skyblue")) +
  labs(title = "Education",
       x = "Education",
       y = "Proportion") +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

# Group and summarize survey_data by state
df_survey <- survey_data %>% 
  count(state) %>%
  mutate(sur_prop = n / sum(n)) %>%
  select(state, sur_prop)

# Group and summarize poststrat_data (census data) by state
df_census <- poststrat_data %>% 
  count(state) %>% 
  mutate(census_prop = n / sum(n)) %>%
  select(state, census_prop)

# Merge the two datasets
combined_data <- left_join(df_census, df_survey, by = "state")

# Plotting
ggplot(combined_data, aes(state)) +
  geom_point(aes(y = census_prop, color = "Census data"), size = 3) +
  geom_point(aes(y = sur_prop, color = "Survey data"), size = 3) +
  scale_color_manual(values = c("Census data" = "salmon", "Survey data" = "skyblue")) +
  labs(title = "State",
       x = "State",
       y = "Proportion") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```
@fig-proportions show a comparison of the survey and census data proportions. We notice that for all the variables of interest, the survey is not entirely representative of the population. For example, we see that females are underrepresented in the survey with the survey having about 49% females whereas the census shows 51% females. We also see how population with highest education attained being high school are under-represented whereas population with highest education attained being more than high school are over-representated. Difference in proportions like these can be seen in all the variables of interests and this is why post-stratifying is important to ensure precision and better representation of the population.

# Model
We used a Bayesian Logistic Regression model with Post-Stratification to predict the 2024 US Presidential Elections. Logistic regression is a method used for binary classification to predict the probability of a categorical dependent variable.

For our analysis, a logistic regression model will be first used to model the proportion of the voters who will vote for Biden. The model will be based on five independent demographic variables: state, sex, age, education and income.

The logistic regression model we will be using is:
\begin{equation}
\log\left(\frac{\hat{p}}{1 - \hat{p}}\right) = \beta_0 + \beta_1 \times \text{state} + \beta_2 \times \text{sex} + \beta_3 \times \text{age} + \beta_4 \times \text{education} + \beta_5 \times \text{income} \label{eq:model1}
\end{equation}

$$
\begin{aligned}
\beta_0 & \sim \mbox{Normal}(0, 2.5)\\
\beta_1 & \sim \mbox{Normal}(0, 2.5)\\
\beta_2 & \sim \mbox{Normal}(0, 2.5)\\
\beta_3 & \sim \mbox{Normal}(0, 2.5)
\end{aligned}
$$

where,

- $\hat{p}$ represents the probability that someone will vote for Biden
- $\beta_0$ represents the intercept term of this logistical regression. It is the probability that someone will vote for Biden if the predictors' values are zero
- $\beta_1$ is the coefficient corresponding to state
- $\beta_2$ is the coefficient corresponding to sex
- $\beta_3$ is the coefficients corresponding to age
- $\beta_4$ is the coefficients corresponding to education
- $\beta_5$ is the coefficients corresponding to income

In our model, normal priors with a mean of 0 and a standard deviation of 2.5 are used for both the coefficients and the intercept. Setting the mean of the priors to 0 implies that there is no expectation of a particular direction or magnitude for the coefficients or intercept. We chose this as we have no expectation of the same. The standard deviation of 2.5 reflects the uncertainty or variability in the prior beliefs. We chose a moderately wide prior to allow for a reasonable amount of uncertainty.

The chosen priors allow the data to largely determine the posterior distribution as they are relatively non-informative. They don't heavily influence the results unless the data provide strong evidence to the contrary.

The use of moderately wide priors can also help regularize the model, preventing overfitting and providing more stable estimates, particularly when dealing with limited data.

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: tbl-model-coefficients
#| tbl-cap: Coefficients of the model

# Create model summary
model_summary <- modelsummary(us_election_model)

# Extract coefficients
coefficients <- broom::tidy(us_election_model, conf.int = TRUE)
coefficients |> kable(digits = 2)
```
@tbl-model-coefficients shows the coefficients for our model that can be used in the equation along with the standard error and the 95% confidence interval. The standard error (SE) is a measure of the precision with which a sample statistic estimates a population parameter. It quantifies the variability of sample statistics around the population parameter. A 95% confidence interval means that if you were to take many samples and construct confidence intervals in the same way, approximately 95% of those intervals would contain the true population mean.

After we created the model, we then post-stratify these proportions and to conclude who we predict to win the 2024 US Presidential Elections. Post-stratification is a statistical technique employed to refine survey estimates by aligning them with known population characteristics. It helps us ensure that our findings accurately represent the diversity of the entire population.

In this process, we first the organize the survey data based on the demographic factors we decided to investigate. This involved counting individuals in each group to understand the sample distribution and computing proportions to depict the relative contribution of each demographic group to the overall survey sample.

Predictive model is then employed to generate probabilities for each demographic group, capturing the nuanced political preferences within distinct strata. These probabilities were summarized by aggregating them within each demographic group.

Applying this post-stratification process results in enhanced predictions that align with the known population distribution within each stratum. These adjusted values contribute to a more accurate representation of party
votes within specific demographic categories, ensuring our findings are reflective of the broader population.

# Results
```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-support-state
#| fig-cap: Estimated proportion support for Biden by State

# Create the plot
ggplot(state_vote_for_biden, aes(y = biden_predict, x = reorder(state, -biden_predict), color = "MRP estimate")) +
  geom_point() +
  geom_errorbar(aes(ymin = biden_lower, ymax = biden_upper), width = 0) +
  geom_hline(yintercept = 0.5, linetype = "dashed") +
  labs(
    x = "State",
    y = "Estimated proportion support for Biden",
    color = "Source"
  ) +
  theme_classic() +
  scale_color_brewer(palette = "Set1") +
  coord_flip() +
  theme(legend.position = "bottom")

```

In @fig-support-state we see the percentage of the votes for Biden/Democratic party. This graph uses the data we
collected after the post-stratification. Here, the red dots on the graph represent the voting percentage and the red lines are the error bars. The black vertical line with dashes indicates the line of proportion. Here, it is the equal split between support for Biden and support for the opposing candidate, Trump.

The states that have points to the left of the black vertical line are the states voting for Trump,
and the states on the right are voting for Biden. We see that the majority of the votes for the democratic party
are found from the Northeast,Midwest  and the West Coast states. For example, the difference in votes for Biden
between Oklahoma and New Mexico is more than 50%, but states like Oregon and Maine are neutral in terms of voting for either parties.

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-support-age
#| fig-cap: Estimated proportion support for Biden by Age Groups

ggplot(age_vote_for_biden, aes(x = age, y = biden_predict)) +
  geom_col(fill = "purple") +
  labs(title = "Proportions of Votes for Biden by age group",
       y = "Proportion of Votes",
       x = "Age groups")

```

In @fig-support-age, we used the data to find the proportion of the voters categorized by age, who would vote for the Democratic party. We see that the younger generation are more likely to vote for Biden compared to the older. Americans under the age of 45 are 70% likely to vote for Trump. The support for Trump drops to around 60% in the 45-59 age group. While this is favorable towards Biden, it does indicate that more people between 45 to 59 years are likely to vote for Trump compared to other age groups.

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-support-sex
#| fig-cap: Estimated proportion support for Biden by Sex

ggplot(sex_vote_for_biden, aes(x = sex, y = biden_predict)) +
  geom_col(fill = "red") +
  labs(title = "Proportions of Votes for Biden by sex",
       y = "Proportion of Votes",
       x = "Sex groups")

```
@fig-support-sex shows the proportion of Female voters that would vote for the Democratic Party as compared to the Male voters. We see a 10% difference in the proportion where women are 10% more likely to vote for Biden.

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-support-general
#| fig-cap: Estimated proportion support for Biden

# Plotting
general_data <- survey_data %>%
  group_by(vote_biden) %>%
  summarise(n = n()) %>%
  mutate(prop = n / sum(n))

ggplot(general_data, aes(x = as.factor(vote_biden), y = prop)) +
  geom_col(fill = "blue") +
  labs(title = "Proportions of Votes",
       y = "Proportion",
       x = "Vote") +
  scale_x_discrete(labels = c("Vote for Trump", "Vote for Biden"))

```

@fig-support-general looks at the division of voters for each party as an overall. We see that even after applying post-stratification, the preference has only changed by a negligible amount with a large majority of the population likely to vote for Trump. This graph clearly shows the conclusion we can deduce from our research that it is clear that the population of the United States is more inclined towards the Democratic Party.

# Appendix

## Cleaning
For the survey dataset, the cleaning steps we took were:

1. Initial cleaning involved filtering the dataset to retain only responses relevant to the analysis. Responses were filtered based on the participant's stance on Q1_1 which is “Overall, do you have a favorable or unfavorable impression of...Donald Trump?”, specifically selecting those labeled as "Favorable" or "Unfavorable".

2. Age groups were defined based on the participants' reported age (ppage). Participants were categorized into four age groups: "18-29", "30-44", "45-59", and "60+". The final was named age.

3. Respondents' income information (ppinc7) was retained in its original format to provide insights into income distribution. The final column was named income.

4. Education levels were standardized to facilitate analysis. Respondents' education information (ppeducat) was categorized into three groups: "High school", "More than high school" (including "Bachelors degree or higher" and "Some college"), and "Less than high school". This was done to match the education level separation available in the poststratification dataset. The final column was named education.

5. The state of residence for each participant (ppstaten) was preserved for regional analysis. The final column was named state.

6. Gender information (ppgender) was just had a change of turning the value to lower case. The final column was named sex as just having these two criterias implies that and to maintain with the poststratification data reporting sex as well.

7. We created a binary variable for favorability of Biden. Responses from Q1_1 labeled as "Favorable" were coded as 0, while responses labeled as "Unfavorable" were coded as 1. The final column was named vote_biden.

8. The final survey dataset contained the above mentioned 6 columns: age, education, income, state, sex, and the coded variable for Biden vote.

For the poststratification dataset, the cleaning steps we took were:

1. Initially, the "AGE" variable was converted to numeric format. Subsequently, data points were filtered to retain only those individuals aged 18 and above, excluding individuals from Vermont and Montana and cases where total household income (FTOTINC) was recorded as less than zero. Individuals from Vermont and Montana were excluded from the dataset to maintain consistency with the survey data as the survey data does not involve any records from these two states.

2. Participants' ages (AGE) were categorized into four groups: "18-29", "30-44", "45-59", and "60+". The final column was named age.

3. Education levels were standardized into four categories: "High school", "More than high school", and "Less than high school". This categorization was based on the reported education level (EDUC), with corresponding numerical codes for each category. The final column was named education.

4. States were labeled according to their corresponding numerical codes (STATEICP), assigning each code to its respective state name. The final column was named education.

5. Sex information (SEX) was changed for the codes to their respective values: "male" and "female". The final column was named sex.

6. Household income (FTOTINC) was categorized into the same division of income brackets as the survey data: Less than $10,000, $10,000 to $24,999, $25,000 to $49,999, $50,000 to $74,999, $75,000 to $99,999, $100,000 to $149,999 and $150,000 or more. The final column was named income.

7. The final poststratification dataset contained the above mentioned 5 columns: age, education, income, state and sex.

Now we are ensured that both the datasets have the same variable names and respective categories.

## Survey dataset
Here is a glimpse of the survey data set used

```{r}
#| echo: false
#| warning: false
#| message: false

head(survey_data)
```

## Poststratification dataset
Here is a glimpse of the poststratification data set used
```{r}
#| echo: false
#| warning: false
#| message: false

head(poststrat_data)
```

# References